---
title: "Survival Analysis Models"
output: html_notebook
---
Three models are included thus far. 

A quatrics model(p1 - Final)

A discrete model (p1-p7 as linear, and p8.optional & Final as dummies)

A regional model (p1-p7, whereas p8 & Final are excluded)

Each of the models contains a base model and a final model.

The base model only includes milestones.

The final model includes all significant predictors and interaction terms.

```{r include=FALSE}
library(rmarkdown)
library(stargazer)
#load(url("https://raw.github.com/bossaround/MOOC_CS50/master/CS50_MOOC_Survival_Analysis_Data_0.RData"))
#use dat9 from Data_1, the time var is milestones, make sure it is plural.
load(url("https://raw.github.com/bossaround/MOOC_CS50/master/CS50_MOOC_Survival_Analysis_Data_1.RData"))

```
```{r include=FALSE}
# dat7$csexperience <- as.numeric(dat7$csexperience)
# dat7$motiv <- as.numeric(dat7$motiv)
dat9$csexperience <- as.numeric(dat9$csexperience)
dat9$motiv <- as.numeric(dat9$motiv)
dat8 <- dat9
dat8$m8 <- 0
dat8$mf <- 0
dat8$m8[which(dat8$milestones==8)] <- 1
dat8$mf[which(dat8$milestones==9)] <- 1
```
```{r include=FALSE}
dat8$milestones[which(dat8$milestones==8)] <- 0
dat8$milestones[which(dat8$milestones==9)] <- 0

```
```{r include=FALSE}
# This is the model used to draw the hazard plot with an 
# interaction effect with male
model.for.plot1 <- glm(drop ~ milestones + milestones2 + milestones3+ male +milestones*male +  age +  testtot  +foreign + gamehour5 + testtot*milestones  , family="binomial", data=dat9)

summary(model.for.plot1)
```
```{r include=FALSE}
m.quatrics.base <- glm(drop ~ milestones + milestones2 + milestones3, family="binomial", data=dat9)
summary(m.quatrics.base)
```

```{r include=FALSE}
m.discrete.base <- glm(drop ~ milestones + m8 + mf, family = "binomial", data = dat8)
summary(m.discrete.base)
```
```{r include=FALSE}
dat10 <- dat9[dat9$milestones <8,]
```

```{r include=FALSE}
m.quatrics.full <- glm(drop ~ milestones +milestones2+csexperience+male+school +testtot +age+motiv+gamehour5+foreign+ milestones*testtot+milestones*csexperience +milestones*male +milestones*age , family = "binomial", data=dat9)
summary(m.quatrics.full)
```

```{r include=FALSE}
m.discrete.full <- glm(drop ~ milestones +m8 + mf +csexperience+male +testtot +age+school+motiv+gamehour5+foreign+ milestones*testtot+ m8*testtot + mf*testtot+milestones*csexperience  + m8*csexperience + mf*csexperience + milestones*male + m8*male + mf*male + milestones*age + m8*age + mf*age , family = "binomial", data=dat8)
summary(m.discrete.full)
```
```{r include=FALSE}
m.regional.base <- glm(drop ~ milestones , family = "binomial", data=dat10)
summary(m.regional.base)
```
```{r include=FALSE}
m.regional.full <- glm(drop ~ milestones +csexperience+male +testtot +age+motiv+gamehour5+foreign+ milestones*testtot+milestones*csexperience +milestones*male + milestones*age, family = "binomial", data=dat9)
summary(m.regional.full)
```
table all models 
```{r results="asis",warning=FALSE,message=FALSE, echo=FALSE}
 stargazer :: stargazer(m.quatrics.base,m.quatrics.full,m.discrete.base,m.discrete.full,m.regional.base,m.regional.full , type="html",
                        dep.var.labels = "Logit Hazard",
                       title = "Survival analysis predicting logit hazard of dropout",
                       model.names = FALSE,
                       column.labels = c("quatrics.\t base","quatrics.\t final","discrete.\t base","discrete.\t final","regional.\t base","regional.\t final") )
```
```{r include=FALSE}
 d.quatrics.complete <- dat9[!is.na(dat9$reason),]
 d.discrete.complete <- dat8[!is.na(dat8$reason),] 
newdat.quatrics <- data.frame(milestones=d.quatrics.complete$milestones, milestones2=d.quatrics.complete$milestones2,milestones3 = d.quatrics.complete$milestone3, male = d.quatrics.complete$male, age = d.quatrics.complete$age,  testtot = d.quatrics.complete$testtot, foreign = d.quatrics.complete$foreign, extrovert = d.quatrics.complete$extrovert, gamehour5 = d.quatrics.complete$gamehour5, csexperience = d.quatrics.complete$csexperience, motiv = d.quatrics.complete$motiv, school=d.quatrics.complete$school)
#predicted log hazard value based on newdat for quatrics
newdat.quatrics$predict_log <- predict(m.quatrics.full, newdat.quatrics, type='link')
#predicted probability based on newdat for quatrics
newdat.quatrics$predict_probability <- predict(m.quatrics.full, newdat.quatrics, type='response')
newdat.quatrics$csexperience.above.mean <- ifelse(newdat.quatrics$csexperience>0,1,0)
newdat.quatrics$testtot.above.mean <- ifelse(newdat.quatrics$testtot>0.1,1,0)

# fitline <- predict(timevaryingmodel2, newdat, type='link',se.fit = TRUE)

# # datK$sefit <- fitline$se.fit
# datK$predict_term <- predict(timevaryingmodel2, newdat, type='term')
 newdat.discrete <- data.frame(milestones=d.discrete.complete$milestones, m8=d.discrete.complete$m8,mf = d.discrete.complete$mf, male = d.discrete.complete$male, age = d.discrete.complete$age,  testtot = d.discrete.complete$testtot, foreign = d.discrete.complete$foreign, extrovert = d.discrete.complete$extrovert, gamehour5 = d.discrete.complete$gamehour5, csexperience = d.discrete.complete$csexperience, motiv = d.discrete.complete$motiv, school=d.quatrics.complete$school)

# #predicted log hazard value based on newdat for discrete
 newdat.discrete$predict_log <- predict(m.discrete.full, newdat.discrete, type='link')
# #predicted probability based on newdat for discrete
 newdat.discrete$predict_probability <- predict(m.discrete.full, newdat.discrete, type='response')
 newdat.discrete$csexperience.above.mean <- ifelse(newdat.quatrics$csexperience>0,1,0)
newdat.discrete$testtot.above.mean <- ifelse(newdat.quatrics$testtot>0.1,1,0)
newdat.discrete.cont <- newdat.discrete
newdat.discrete.cont$milestones[which(newdat.discrete.cont$m8==1)] <- 8
newdat.discrete.cont$milestones[which(newdat.discrete.cont$mf==1)] <- 9

```
Draw plot

```{r warning=FALSE}
labs <- c("p1", "p2", "p3", "p4","p5","p6","p7","p8 (optional)","Final")

 ggplot()+
  geom_smooth (data= newdat.quatrics[newdat.quatrics$milestones<8,], method="glm", formula = y~x,  alpha=0.65, size = 1,level=0.99, aes(milestones, predict_probability,  fill=as.factor(csexperience.above.mean) , color=as.factor(csexperience.above.mean))) +
  geom_smooth (data=newdat.quatrics, method="glm", formula = y~x+I(x^2),  alpha=0.65, size = 1,level=0.99, aes(milestones, predict_probability,  fill=as.factor(csexperience.above.mean) , color=as.factor(csexperience.above.mean)))
```
```{r warning=FALSE}
labs <- c("p1", "p2", "p3", "p4","p5","p6","p7","p8 (optional)","Final")

 ggplot()+
  geom_smooth (data= newdat.discrete[newdat.discrete$milestones>0,], method="glm", formula = y~x,  alpha=0.65, size = 1,level=0.99, aes(milestones, predict_probability,  fill=as.factor(testtot.above.mean) , color=as.factor(testtot.above.mean))) +
  geom_smooth (data=newdat.discrete.cont, method="glm", formula = y~x+I(x^2),  alpha=0.65, size = 1,level=0.99, aes(milestones, predict_probability,  fill=as.factor(testtot.above.mean) , color=as.factor(testtot.above.mean))) +
   geom_pointrange(data=newdat.discrete.cont[newdat.discrete.cont$m8==1,],aes(milestones,mean(predict_probability), ymin=mean(predict_probability)-0.06, ymax=mean(predict_probability)+0.06, color=as.factor(testtot.above.mean)))
```
```{r warning=FALSE}
labs <- c("p1", "p2", "p3", "p4","p5","p6","p7","p8 (optional)","Final")

 ggplot()+
  geom_smooth (data= newdat.quatrics[newdat.quatrics$milestones<8,], method="glm", formula = y~x,  alpha=0.65, size = 1,level=0.99, aes(milestones, predict_probability,  fill=as.factor(male) , color=as.factor(male))) +
  geom_smooth (data=newdat.quatrics, method="glm", formula = y~x+I(x^2),  alpha=0.65, size = 1,level=0.95, aes(milestones, predict_probability,  fill=as.factor(male) , color=as.factor(male)))
```

