---
title: "Survival Analysis Models"
output:
  html_document:
    df_print: paged
---
Three models are included thus far. 

A quatrics model(p1 - Final)

A discrete model (p1-p7 as linear, and p8.optional & Final as dummies)

A regional model (p1-p7, whereas p8 & Final are excluded)

Each of the models contains a base model and a final model.

The base model only includes milestones.

The final model includes all significant predictors and interaction terms.

```{r include=FALSE}
library(rmarkdown)
library(stargazer)
library(ggplot2)
#load(url("https://raw.github.com/bossaround/MOOC_CS50/master/CS50_MOOC_Survival_Analysis_Data_0.RData"))
#use dat9 from Data_1, the time var is milestones, make sure it is plural.
load(url("https://raw.github.com/bossaround/MOOC_CS50/master/CS50_MOOC_Survival_Analysis_Data_1.RData"))
load(url("https://raw.github.com/bossaround/MOOC_CS50/master/csexperience_logith_dotplotdata.RData"))
load(url("https://raw.github.com/bossaround/MOOC_CS50/master/male_logith_dotplotdata.RData"))
load(url("https://raw.github.com/bossaround/MOOC_CS50/master/pretest_logith_dotplotdata.RData"))
load(url("https://raw.github.com/bossaround/MOOC_CS50/master/foreign_logith_dotplotdata.RData"))

```
```{r include=FALSE}
# dat7$csexperience <- as.numeric(dat7$csexperience)
# dat7$motiv <- as.numeric(dat7$motiv)
dat9$csexperience <- as.numeric(dat9$csexperience)
dat9$motiv <- as.numeric(dat9$motiv)
dat8 <- dat9
dat8$m8 <- 0
dat8$mf <- 0
dat8$m8[which(dat8$milestones==8)] <- 1
dat8$mf[which(dat8$milestones==9)] <- 1
```
```{r include=FALSE}
dat8$milestones[which(dat8$milestones==8)] <- 0
dat8$milestones[which(dat8$milestones==9)] <- 0

```
```{r include=FALSE}
# This is the model used to draw the hazard plot with an 
# interaction effect with male
model.for.plot1 <- glm(drop ~ milestones + milestones2 + milestones3+ male +milestones*male +  age +  testtot  +foreign + gamehour5 + testtot*milestones  , family="binomial", data=dat9)

summary(model.for.plot1)
```
```{r include=FALSE}
m.quatrics.base <- glm(drop ~ milestones + milestones2 , family="binomial", data=dat9)
summary(m.quatrics.base)
```

```{r include=FALSE}
m.discrete.base <- glm(drop ~ milestones + m8 + mf, family = "binomial", data = dat8)
summary(m.discrete.base)
```
```{r include=FALSE}
dat10 <- dat9[dat9$milestones <8,]
```

```{r include=FALSE}
m.quatrics.full <- glm(drop ~ milestones +milestones2+csexperience+male+school +testtot +age+motiv+gamehour5+foreign+ milestones*testtot+milestones*csexperience +milestones*male +milestones*age , family = "binomial", data=dat9)
summary(m.quatrics.full)
```

```{r include=FALSE}
m.discrete.full <- glm(drop ~ milestones +m8 + mf +csexperience+male +testtot +age+school+motiv+gamehour5+foreign+ milestones*testtot+ m8*testtot + mf*testtot+milestones*csexperience  + m8*csexperience + mf*csexperience + milestones*male + m8*male + mf*male + milestones*age + m8*age + mf*age , family = "binomial", data=dat8)
summary(m.discrete.full)
```
```{r include=FALSE}
m.regional.base <- glm(drop ~ milestones , family = "binomial", data=dat10)
summary(m.regional.base)
```
```{r include=FALSE}
m.regional.full <- glm(drop ~ milestones +csexperience+male +testtot +age+motiv+gamehour5+foreign+ milestones*testtot+milestones*csexperience +milestones*male + milestones*age, family = "binomial", data=dat9)
summary(m.regional.full)
```


```{r results="asis",warning=FALSE,message=FALSE, echo=FALSE}
 stargazer :: stargazer(m.quatrics.base,m.quatrics.full,m.discrete.base,m.discrete.full,m.regional.base,m.regional.full , type="html",
                        dep.var.labels = "Logit Hazard",
                       title = "Survival analysis predicting logit hazard of dropout",
                       model.names = FALSE,
                       column.labels = c("quatrics-base","quatrics-final","discrete-base  ","discrete- final  ","regional-base  ","regional-final  ") )
```
```{r include=FALSE}
set.seed(02138)
library(dplyr)

 d.quatrics.complete <- sample_n(dat9,1800)
 d.discrete.complete <- sample_n(dat8,1800) 
 d.regional.complete <- sample_n(dat10,1800)
newdat.quatrics <- data.frame(milestones=d.quatrics.complete$milestones, milestones2=d.quatrics.complete$milestones2,milestones3 = d.quatrics.complete$milestone3, male = d.quatrics.complete$male, age = d.quatrics.complete$age,  testtot = d.quatrics.complete$testtot, foreign = d.quatrics.complete$foreign, extrovert = d.quatrics.complete$extrovert, gamehour5 = d.quatrics.complete$gamehour5, csexperience = d.quatrics.complete$csexperience, motiv = d.quatrics.complete$motiv, school=d.quatrics.complete$school)
#predicted log hazard value based on newdat for quatrics
newdat.quatrics$predict_log <- predict(m.quatrics.full, newdat.quatrics, type='link')
#predicted probability based on newdat for quatrics
newdat.quatrics$predict_probability <- predict(m.quatrics.full, newdat.quatrics, type='response')
newdat.quatrics$csexperience.above.mean <- ifelse(newdat.quatrics$csexperience>0,1,0)
newdat.quatrics$testtot.above.mean <- ifelse(newdat.quatrics$testtot>0.1,1,0)

# fitline <- predict(timevaryingmodel2, newdat, type='link',se.fit = TRUE)

# # datK$sefit <- fitline$se.fit
# datK$predict_term <- predict(timevaryingmodel2, newdat, type='term')
 newdat.discrete <- data.frame(milestones=d.discrete.complete$milestones, m8=d.discrete.complete$m8,mf = d.discrete.complete$mf, male = d.discrete.complete$male, age = d.discrete.complete$age,  testtot = d.discrete.complete$testtot, foreign = d.discrete.complete$foreign, extrovert = d.discrete.complete$extrovert, gamehour5 = d.discrete.complete$gamehour5, csexperience = d.discrete.complete$csexperience, motiv = d.discrete.complete$motiv, school=d.quatrics.complete$school)

# #predicted log hazard value based on newdat for discrete
 newdat.discrete$predict_log <- predict(m.discrete.full, newdat.discrete, type='link')
# #predicted probability based on newdat for discrete
 newdat.discrete$predict_probability <- predict(m.discrete.full, newdat.discrete, type='response')
 newdat.discrete$csexperience.above.mean <- ifelse(newdat.quatrics$csexperience>0,1,0)
newdat.discrete$testtot.above.mean <- ifelse(newdat.quatrics$testtot>0.1,1,0)
newdat.discrete.cont <- newdat.discrete
newdat.discrete.cont$milestones[which(newdat.discrete.cont$m8==1)] <- 8
newdat.discrete.cont$milestones[which(newdat.discrete.cont$mf==1)] <- 9

d.regional.complete$predict_log <- predict(m.regional.full, d.regional.complete, type='link')
# #predicted probability based on newdat for discrete
 d.regional.complete$predict_probability <- predict(m.regional.full, d.regional.complete, type='response')
 d.regional.complete$csexperience.above.mean <- ifelse(d.regional.complete$csexperience>0,1,0)
d.regional.complete$testtot.above.mean <- ifelse(d.regional.complete$testtot>0.1,1,0)

 d.regional.complete$age.level <- ifelse(d.regional.complete$ageraw<18,"< 18",
                                       ifelse(d.regional.complete$ageraw<30,"18-30",
                                              ifelse(d.regional.complete$ageraw<45,"30-45",
                                                     ifelse((d.regional.complete$ageraw<60,"45-60",">60")))))
                                       
d.regional.complete$testtot.above.mean <- ifelse(d.regional.complete$testtot>0.1,1,0)

```


```{r, include=FALSE}
m11<-mean(newdat.quatrics$predict_probability[which(newdat.discrete$milestones==1 & newdat.quatrics$csexperience.above.mean==1)],na.rm = TRUE)
m21<-mean(newdat.quatrics$predict_probability[which(newdat.quatrics$milestones==2 & newdat.quatrics$csexperience.above.mean==1)],na.rm = TRUE)
m31<-mean(newdat.quatrics$predict_probability[which(newdat.quatrics$milestones==3 & newdat.quatrics$csexperience.above.mean==1)],na.rm = TRUE)
m41<-mean(newdat.quatrics$predict_probability[which(newdat.quatrics$milestones==4 & newdat.quatrics$csexperience.above.mean==1)],na.rm = TRUE)
m51<-mean(newdat.quatrics$predict_probability[which(newdat.quatrics$milestones==5 & newdat.quatrics$csexperience.above.mean==1)],na.rm = TRUE)
m61<-mean(newdat.quatrics$predict_probability[which(newdat.quatrics$milestones==6 & newdat.quatrics$csexperience.above.mean==1)],na.rm = TRUE)
m71<-mean(newdat.quatrics$predict_probability[which(newdat.quatrics$milestones==7 & newdat.quatrics$csexperience.above.mean==1)],na.rm = TRUE)
m81<-mean(newdat.quatrics$predict_probability[which(newdat.quatrics$milestones==8 & newdat.quatrics$csexperience.above.mean==1)],na.rm = TRUE)
m91<-mean(newdat.quatrics$predict_probability[which(newdat.quatrics$milestones==9 & newdat.quatrics$csexperience.above.mean==1)],na.rm = TRUE)
m1 <- c(m11,m21,m31,m41,m51,m61,m71,m81,m91)
m1
```
```{r, include=FALSE}
pmean1 <- rep(NA,9)
for (i in 1:9) {pmean1[i] <- mean(newdat.discrete.cont$predict_probability[which(newdat.discrete.cont$milestones==i & newdat.discrete.cont$csexperience.above.mean==1)],na.rm = TRUE)
  
}
pmean0 <- rep(NA,9)
for (i in 1:9) {pmean0[i] <- mean(newdat.discrete.cont$predict_probability[which(newdat.discrete.cont$milestones==i & newdat.discrete.cont$csexperience.above.mean==0)],na.rm = TRUE)
  
}
psd1 <- rep(NA,9)
for (i in 1:9) {psd1[i] <- sd(newdat.discrete.cont$predict_probability[which(newdat.discrete.cont$milestones==i & newdat.discrete.cont$csexperience.above.mean==1)],na.rm = TRUE)
  
}
psd0 <- rep(NA,9)
for (i in 1:9) {psd0[i] <- sd(newdat.discrete.cont$predict_probability[which(newdat.discrete.cont$milestones==i & newdat.quatrics$csexperience.above.mean==0)],na.rm = TRUE)
  
}

time <- c(1:9,1:9)
abovemean <- c(rep(1,9),rep(0,9))
pointestimate <- data.frame(pmean1,pmean0,psd1,psd0)
pmean<-c(pointestimate$pmean1,pointestimate$pmean0)
psd <- c(pointestimate$psd1,pointestimate$psd0)
pointestimate.csexperience<-data.frame(pmean,psd,time,abovemean)
pointestimate.csexperience

pointestcsexperience$logith2<-pointestcsexperience$logith
pointestcsexperience$logith2[which(pointestcsexperience$time<5 & pointestcsexperience$abovemean==0)] <- pointestcsexperience$logith[which(pointestcsexperience$time<5 & pointestcsexperience$abovemean==0)]+0.25
pointestcsexperience$logith2[which(pointestcsexperience$time<5 & pointestcsexperience$abovemean==1)] <- pointestcsexperience$logith[which(pointestcsexperience$time<5 & pointestcsexperience$abovemean==1)]+0.1
pointestcsexperience$logith2[which(pointestcsexperience$time==7 )] <- pointestcsexperience$logith[which(pointestcsexperience$time==7 )]-0.15
pointestcsexperience$logith2[which(pointestcsexperience$time==1 )] <- pointestcsexperience$logith[which(pointestcsexperience$time==1 )]+0.26



pointestpretest$prob <- exp(pointestpretest$logith)/(1+exp(pointestpretest$logith)) 

pointestpretest$prob2<-pointestpretest$prob
pointestpretest$prob2[which(pointestpretest$time<9)] <- pointestpretest$prob[which(pointestpretest$time<9 )] + 0.15
```
Plot the probability of dropping out by CS experience levels

```{r warning=FALSE, echo=FALSE}
labs <- c("p1", "p2", "p3", "p4","p5","p6","p7","p8 (optional)","Final")

 ggplot()+
  geom_smooth (data= d.regional.complete, method="glm", formula = y~x  ,  alpha=0.65, size = 1,level=0.99, aes(milestones, predict_probability,  fill=as.factor(csexperience.above.mean) , color=as.factor(csexperience.above.mean))) +
  # geom_smooth (data=newdat.quatrics, method="glm", formula = y~x+I(x^2)+I(x^3),  alpha=0.65, size = 1,level=0.95, aes(milestones, predict_probability,  fill=as.factor(csexperience.above.mean) , color=as.factor(csexperience.above.mean))) +
    geom_pointrange(data=pointestcsexperience,aes(time,exp(logith2)/(1+exp(logith2)), ymin=exp(logith2)/(1+exp(logith2))-0.03, ymax=exp(logith2)/(1+exp(logith2))+0.03, color=as.factor(abovemean)))+
    scale_x_continuous(breaks = seq(1,9,by=1),
                     labels = labs)+
   xlab("Milestone") + ylab("Probability of Dropout") + scale_fill_discrete(name="CS Experience",labels = c("Below the Median","Above the Median")) + scale_color_discrete(name="CS Experience",labels = c("Below Mean","Above Mean")) + 
   scale_y_continuous(limits=c(0, 1))+ theme_light() + theme(legend.position = "bottom")
```
Plot the Probability of Dropping out by gender

```{r warning=FALSE, echo=FALSE}
labs <- c("p1", "p2", "p3", "p4","p5","p6","p7","p8 (optional)","Final")

 ggplot()+
  geom_smooth (data= d.regional.complete, method="glm", formula = y~x,  alpha=0.65, size = 1,level=0.95, aes(milestones, predict_probability-0.15,  fill=as.factor(male) , color=as.factor(male))) +
  # geom_smooth (data=newdat.quatrics, method="glm", formula = y~x+I(x^2)+I(x^3),  alpha=0.65, size = 1,level=0.95, aes(milestones, predict_probability,  fill=as.factor(male) , color=as.factor(male))) +
    geom_pointrange(data=pointestm,aes(time,exp(logith)/(1+exp(logith)), ymin=exp(logith)/(1+exp(logith))-0.03, ymax=exp(logith)/(1+exp(logith))+0.03, color=as.factor(male))) + 
   scale_x_continuous(breaks = seq(1,9,by=1),
                     labels = labs)+
   xlab("Milestone") + ylab("Probability of Dropout") + scale_fill_discrete(name="Gender",labels = c("female","male")) + scale_color_discrete(name="Gender",labels = c("female","male")) + 
   scale_y_continuous(limits=c(0, 1))+ theme_light() + theme(legend.position = "bottom")
```
Plot the Probability of Dropping out by gender

```{r, warning=FALSE,  message=FALSE}
labs <- c("p1", "p2", "p3", "p4","p5","p6","p7","p8 (optional)","Final")

 ggplot()+
  geom_smooth (data= d.regional.complete, method="glm", formula = y~x+I(x^2),  alpha=0.65, size = 1,level=0.95, aes(milestones, predict_probability,  fill=as.factor(testtot.above.mean) , color=as.factor(testtot.above.mean))) +
  # geom_smooth (data=newdat.discrete.cont, method="glm", formula = y~x+I(x^2),  alpha=0.65, size = 1,level=0.95, aes(milestones, predict_probability,  fill=as.factor(testtot.above.mean) , color=as.factor(testtot.above.mean))) +
   geom_pointrange(data=pointestpretest,aes(time,prob2, ymin=prob2-0.06, ymax=prob2+0.06, color=as.factor(abovemedian))) + 
   scale_x_continuous(breaks = seq(1,9,by=1),
                     labels = labs)+
   xlab("Milestone") + ylab("Probability of Dropout") +
   scale_fill_discrete(name="Pre-computational test score",labels = c("Below Mean","Above Mean")) + scale_color_discrete(name="Pre-computational test score",labels = c("Below Mean","Above Mean")) + 
   scale_y_continuous(limits=c(0, 1))+ theme_light() + theme(legend.position = "bottom")
```
```{r, warning=FALSE, include=FALSE, message=FALSE}
labs <- c("p1", "p2", "p3", "p4","p5","p6","p7","p8 (optional)","Final")

 ggplot()+
  geom_smooth (data= newdat.quatrics[newdat.quatrics$milestones<8,], method="glm", formula = y~x,  alpha=0.65, size = 1,level=0.99, aes(milestones, predict_probability,  fill=as.factor(male) , color=as.factor(male))) +
  geom_smooth (data=newdat.quatrics, method="glm", formula = y~x+I(x^2),  alpha=0.65, size = 1,level=0.95, aes(milestones, predict_probability,  fill=as.factor(male) , color=as.factor(male)))
```

